---
title: "Analysis on ToothGrowth Data"
author: "Yanhui Zhou"
date: "October 23, 2015"
output: pdf_document
---
# Overview
This is the second portion of the statistical inference class' course project of Coursera, we're going to do basic inferential data analysis on the ToothGrowth data in the R datasets package.

## Data Background
The ToothGrowth data is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).  

# Load the ToothGrowth data and exploratory data
Load the ToothGrowth data and show basic summary
```{r}
library(datasets)
data("ToothGrowth")
summary(ToothGrowth)
```  
Plot a figure to show the feature of the data  
```{r}
library(ggplot2)
g <- ggplot(data=ToothGrowth, aes(dose, len, group = supp))
g <- g + facet_grid(facets = .~supp)
g <- g + geom_point(aes(colour = dose))
g <- g + geom_smooth(stat = "smooth", method = "loess")
suppressWarnings(print(g))
```

# analyze intervals
Below, use one side confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.  
**T test on different supp, Supplement type (VC or OJ)**    
null hypothesis: mean of supplement with "VC" >= mean of supplement with "OJ"  
alternative hypothesis: mean of supplement with "VC" < mean of supplement with "OJ"
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == 'VC'], ToothGrowth$len[ToothGrowth$supp == 'OJ'], 
    var.equal = FALSE, paired = FALSE, alternative="less")
```  
*Result: p-value < 0.05*  
**T test on different Dose(milligrams)**  
1. 0.5 vs. 1.0  
null hypothesis: mean of dose with 0.5 milligrams >= mean of dose with 1.0 milligrams  
alternative hypothesis: mean of dose with 0.5 milligrams < mean of dose with 1.0 milligrams
```{r}
t.test(ToothGrowth$len[ToothGrowth$dos == 0.5], ToothGrowth$len[ToothGrowth$dos == 1.0], 
    var.equal = FALSE, paired = FALSE, alternative="less")
```     
*Result: p-value < 0.05*

2. 1.0 vs. 2.0  
null hypothesis: mean of dose with 1.0 milligrams >= mean of dose with 2.0 milligrams  
alternative hypothesis: mean of dose with 1.0 milligrams < mean of dose with 2.0 milligrams
```{r}
t.test(ToothGrowth$len[ToothGrowth$dos == 1.0], ToothGrowth$len[ToothGrowth$dos == 2.0], 
    var.equal = FALSE, paired = FALSE, alternative="less")
```     
*Result: p-value < 0.05*

# Conclusion
* Assumptions
    1. Suppose the guinea pigs are chose randomly.
    2. The each group of 10 guinea pigs is independent.
    3. Assume a different variance per group.
    
* Conclusion
According to the T test result of null hypothesis and alternative hypothesis, we konw:
    1. Supplement with VC is not better to length of teeth than OJ. So, orange juice(OJ) is more effective.
    2. Vitamin is effective to teeth, as dose increase, length of teeth will be longer.

