---
title: "Exponential Distribution vs. CLT"
author: "zhouyanhui"
date: "October 22, 2015"
output: pdf_document
---

# Overview
This project is for the statistical inference class of Coursera. We will investigate the exponential distribution in R and compare it with the Central Limit Theorem. We will use simulation to explore inference and do some simple inferential data analysis.

# Simulations
Simulate by 40 exponentials with 1000 times, draw the distribution of the averages of 40 exponentials. Then investigate properites of this sample mean distribution, and compare with properites of theoretical distribution ( Central Limit Theorem ) .
```{r}
# set lambda and other constants
lambda <- 0.2
n <- 40
exnum <- 1000

# begin simulation
set.seed(333)
dmatrix <- replicate(exnum, rexp(n, lambda))
mns <- apply(dmatrix, 2, mean)
h<-hist(mns, breaks = 30, prob=TRUE, col='skyblue', 
        xlab='Mean of 40 exponentials', ylab = 'Density', 
        main="Comparision between Sample and Theoretical Distribution",
        cex.main=0.8, cex.lab=0.8)
rug(mns)
# add the line of sample'mean
sample_mean <- mean(mns)
sample_sd <- sd(mns)
abline(v=sample_mean, col='red', lwd = 2)
# add the line of distribution'mean
theory_mean = 1/lambda
theory_sd = 1/lambda * 1/sqrt(n)
abline(v=theory_mean, col='black', lwd = 2)
legend("topright", col=c('red', 'black'), 
       legend=c(paste('sample mean:', round(sample_mean, 2)), 
                paste('theoretical mean:',theory_mean)),
       cex=0.7, lwd=2, box.lwd=0)

# simulate the distribution
xfit<-seq(min(mns),max(mns),length=length(mns)) 
yfit<-dnorm(xfit,mean=sample_mean,sd=sample_sd) 
lines(xfit, yfit, col="green", lwd=5)

tyfit<-dnorm(xfit,mean=theory_mean,sd=theory_sd)
lines(xfit, tyfit, col="violet", lwd=5)

legend("right", col=c('green', 'violet'), 
       legend=c('sample distribution', 'theoretical distribution'), 
       cex=0.7, lwd=5, box.lwd=0)
```

# Sample Mean versus Theoretical Mean, Sample Variance versus Theoretical Variance
In the figure "Comparision between Sample and Theoretical Distribution", the vertical line of sample mean is colored by red, and the theoretical mean is black. The differece is very small. The green curve is the sample distribution and the violet curve is the theoretical distribution of mean of 40 exponentials. And the detail value of means and variances are in the table above.

From the analysis and simulation above, we know, the detail comparision are:

|  Variable  | Theoretical    |   Sample        |
|:-----------|---------------:|:---------------:|
| Mean       |`r theory_mean` |`r sample_mean`  |      
| SD         |`r theory_sd`   |`r sample_sd`    | 
| Variance   |`r theory_sd^2` |`r sample_sd^2`  | 

## Comparision variance of 40 exponentials with theoretical variance
Plot variance of 40 exponentials, to compare with theoretical variance.
```{r}
library(ggplot2)
# begin simulation, using the same date as above
vardat <- data.frame(
    x <- apply(dmatrix, 2, var)
)
ori_theory_var <- (1/lambda)^2
g <- ggplot(vardat, aes(x = x, fill = n)) + geom_density(size = 2, alpha = .2)
g <- g + geom_vline(xintercept = ori_theory_var, size = 2, col='red')
g <- g + geom_text(aes(x = ori_theory_var + 3, y = 0.042,
                       label='theoretical varicane = 25'),
                   hjust =0, vjust =0, colour = 'red')
g <- g + ggtitle('Sample Variance versus Theoretical Variance')
g <- g + xlab('Variance of 40 exponentials')
print(g)
```

From the figure above, we kown the variance of 40 exponentials is also aroud the


# Distribution
From simulation figure "Comparision between Sample and Theoretical Distribution",
we kown the distribution of average of 40 exponentials is approximately normal, with mean=`r sample_mean` close to `r theory_mean`, variance=`r sample_sd^2` close to `r theory_sd^2`. According to Central Limit Theorem, the distribution of averages of iid variables becomes that of a normal with large sample size, approximately $\bar X_n$ ~ $N(\mu, \sigma^2 / n)$.
