---
title: "Regression Models Course Project"
author: "zhouyanhui"
date: "November 19, 2015"
output: pdf_document
---
# Executive Summary  
This analysis is to explore the mtcars data, and try to answer the following two questions:Is an automatic or manual transmission better for MPG? Quantify the MPG difference between automatic and manual transmissions?  
After linear regression analysis, there is a difference between automatic and manual transmission. Manual transmission is better than automatic for MPG. It increases 1.8 MPG swithing from automatic transmission to manual transmission with others not changed.  

# Read Data and Preprocess
```{r}
library(datasets)
data("mtcars")
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am, labels = c("automatic", "manual"))
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```  
# Exploratory Data  
Make a boxplot to show the relation between mpg and am(Transmission), appendix figure 1. 
According to figure 1, mpg of automatic is less than mpg of manual.   

# Regression Model and Model Selection  
Use step function to select the best linear regression model  
```{r}
model <- lm(mpg ~ ., data = mtcars)
bestModel <- step(model, trace=FALSE)
summary(bestModel)
```  
Compare the selected model with the base model and full Model.
```{r}
baseModel <- lm(mpg ~ am, data = mtcars)
fullModel <- lm(mpg ~., data = mtcars)
anova(baseModel, bestModel)$P
anova(bestModel, fullModel)$P
```  
According to the p-value, the bestModel has the most necessary variance to explain outcome. The R-square of bestModel is `r summary(bestModel)$r.square`, that means it can explain `r summary(bestModel)$r.square * 100`% of the origin variance, so the bestModel is significant.

# Residual Analysis  
From Appendix Figure 2. The "Residuals vs Fitted" shows residuals are randomly scattered with no obvious pattern. In "Normal Q-Q", the theoretical quantiles and standardized residuals are almost in a line.

# Conclusion  
1. Comparing to automatic transmission, manual transmission has about 1.8 more miles per gallon.
2. MPG will decrease by about 2.5 for increasing every 1000 lb weight.
3. From 4 cylinders to 6 and 8 cylinders, MPG will decrease about 3.0 and 2.2.
4. Increasing 1 gross horsepower, MPG will only deacrease 0.032.

# Appendix
## Figure 1:  
```{r, echo=FALSE}
suppressMessages(library(ggplot2))
g <- ggplot(mtcars, aes(am, mpg, color = am))
g <- g + geom_point() + geom_boxplot() 
g <- g + labs(title="Miles per gallon with Different Transmission", x='Transmission', y = 'mpg')
print(g)
```  

## Figure 2:
```{r, echo=FALSE}
par(mfrow=c(2, 2))
plot(bestModel)
```
